Online Laterality Battery Pre-Processing - short version

### 2020-06-09 battery-preprocessing_v5_DB.Rmd
updates for logRT

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated May 21st 2020 (now notebook format)"
output: html_notebook
---
23 May 2020: Created subject number for those included in final analysis. CF outlier detection needed updating to ensure only correct trials used for outliers - Correct had not been coded at that point in prior version
22 May 2020: AP made changes to N trials needed for inclusion  
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

@@

demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactor=F)

@@

```{r sightednesss, warning=FALSE}
porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
miles_dat <-  read.csv("https://osf.io/ekqsp//download",stringsAsFactors = F)

@@

EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)

@@

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "https://osf.io/mhk29/download" #day 1 data

  if (d ==2){
    data_file <-  "https://osf.io/ebm6s/download"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

@@

# w<-which(is.na(chim_dat1$subject))
# if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
# w<-which(is.na(chim_dat2$subject))
# if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}

@@

outliertable1<-table(RDT_dat1$subject,RDT_dat1$outlier)
outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)

@@

@@

Comparison shows that there is significant difference in RT. Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
```{r rdtpirate}
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat <-origdat[origdat$side %in% c('left','right'),]
measure <- 'Accuracy'
origdat$X1 <- origdat$p.corr
task <- 'Rhyme Detection'

dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
}
```

```{r RDT.LI}
#now create LI
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"Day1.RDT_p.corr_LI"
if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

#t.test to test if laterality index differs from zero
partsum <- allsum[allsum$exRDT==0,]
print('Testing for difference of single mean from zero')
t.test(partsum$Day1.RDT_p.corr_LI)
t.test(partsum$Day1.RDT_RT_LI)
t.test(partsum$Day2.RDT_p.corr_LI)
t.test(partsum$Day2.RDT_RT_LI)

cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,9,paste0('r = ',round(cor2,3)))
```
## Does RDT relate to handedness?
Answer is convincing nope!

```{r RDThand}
t.test(Day1.RDT_RT_LI~handpch,data=partsum)
t.test(Day2.RDT_RT_LI~handpch,data=partsum)
```
### T-test as alternative to LI for RDT

@@

  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$subject),
                       FUN=mean,na.rm=TRUE)
  colnames(aggdata)<-c('side','item','subject','RT')
  mysubs<-unique(aggdata$subject)
  if(d==1){
  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
  }
  nsubs<-length(mysubs)
  for (i in 1:nsubs){
    myt<-NA
    myp<-NA
    thisdat <- aggdata[aggdata$subject==mysubs[i],]
    pairedt$subject[i]<- mysubs[i]
    #check there are data for both left and right on all items
    checkdat <-table(thisdat$item)
    w<-which(checkdat<2)
    if(length(w)==0){
      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
      myt<-t$statistic
      myp<-t$p.value
    }
    if(d==1){
      pairedt$t1[i] <- myt
      pairedt$p1[i] <-myp
    }
    if(d==2){
      pairedt$t2[i] <-myt
      pairedt$p2[i] <-myp
    }
  }
}
cor(pairedt$t1,pairedt$t2,use='complete.obs')
```

@@

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

>>>>

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated May 21st 2020 (now notebook format)"
output:
  pdf_document: default
  html_notebook: default
---

## Updates

09 June 2020: AP made changes to calcultaions of rhyme
(1) Added t scores to output
(2) computed LIs for log-RT
(3) changed polariies so that a negative value was indicative of left lateralisation in the brain.

23 May 2020: Created subject number for those included in final analysis. CF outlier detection needed updating to ensure only correct trials used for outliers - Correct had not been coded at that point in prior version

22 May 2020: AP made changes to N trials needed for inclusion  
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

@@

demo_dat <- read.csv('./data/Demographics.csv',stringsAsFactor=F)

@@

```{r sightednesss, warning=FALSE}
porta_dat <- read.csv("./data/Porta.csv",stringsAsFactors = F)
miles_dat <-  read.csv("./data/Miles.csv",stringsAsFactors = F)

@@

EHI <- read.csv("./data/EHI.csv",stringsAsFactors = F)

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE <- read.csv("./data/LexTALE.csv",stringsAsFactors = F)

@@

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./data/combined/day1_combined.csv" #day 1 data

  if (d ==2){
    data_file <-  "./data/combined/day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

@@

# w<-which(is.na(chim_dat1$subject))
# if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
# w<-which(is.na(chim_dat2$subject))
# if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
```

@@

outliertable1<-table(RDT_dat1$subject,RDT_dat1$outlier)
outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)
```

@@

# remove NaN
is.nan.data.frame <- function(x)
  do.call(cbind, lapply(x, is.nan))
RDT_df1[is.nan(RDT_df1)] <- NA
RDT_df2[is.nan(RDT_df2)] <- NA

# add log RT
RDT_df1$logRT <- log10(RDT_df1$RT)
RDT_df2$logRT <- log10(RDT_df2$RT)

@@

Comparison shows that there is significant difference in RT. Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
```{r rdtpirate}
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat <-origdat[origdat$side %in% c('left','right'),]

# accuracy
measure <- 'Accuracy'
origdat$X1 <- origdat$p.corr
task <- 'Rhyme Detection'
dopirate(origdat,task,measure)

# RT
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)

# log RT
origdat$X1 <- origdat$logRT
measure <- 'logRT'
dopirate(origdat,task,measure)
}
```

```{r RDT.LI}
#now create LI
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}

# accuracy
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"Day1.RDT_p.corr_LI"
if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# RT
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# log-RT
origdat$X1 <- origdat$logRT
mycolname<-"Day1.RDT_logRT_LI"
if(d==2){mycolname<-"Day2.RDT_logRT_LI"}
polarity <- (-1)
mycutoff <- 0
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

#t.test to test if laterality index differs from zero
partsum <- allsum[allsum$exRDT==0,]
print('Testing for difference of single mean from zero')
t.test(partsum$Day1.RDT_p.corr_LI)
t.test(partsum$Day1.RDT_RT_LI)
t.test(partsum$Day1.RDT_logRT_LI)
t.test(partsum$Day2.RDT_p.corr_LI)
t.test(partsum$Day2.RDT_RT_LI)
t.test(partsum$Day2.RDT_logRT_LI)

# correlate
cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
cor3<-cor(partsum$Day1.RDT_logRT_LI,partsum$Day2.RDT_logRT_LI,use='complete.obs')
# acc plot
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
# raw plot
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,9,paste0('r = ',round(cor2,3)))
# log plot
plot(partsum$Day1.RDT_logRT_LI,partsum$Day2.RDT_logRT_LI,main='Rhyme detect: LI for log-RT')
abline(v=0)
abline(h=0)
text(2,1,paste0('r = ',round(cor3,3)))
```
## Does RDT relate to handedness?
Answer is convincing nope!

```{r RDThand}
t.test(Day1.RDT_RT_LI~handpch,data=partsum)
t.test(Day2.RDT_RT_LI~handpch,data=partsum)
```

### T-test as alternative to LI for RDT

@@

  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$subject),
                       FUN=mean,na.rm=TRUE)
  colnames(aggdata)<-c('side','item','subject','RT')
  mysubs<-unique(aggdata$subject)
  if(d==1){
  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
  }
  nsubs<-length(mysubs)
  for (i in 1:nsubs){
    myt<-NA
    myp<-NA
    thisdat <- aggdata[aggdata$subject==mysubs[i],]
    pairedt$subject[i]<- mysubs[i]
    #check there are data for both left and right on all items
    checkdat <-table(thisdat$item)
    w<-which(checkdat<2)
    if(length(w)==0){
      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
      myt<-t$statistic
      myp<-t$p.value
    }
    if(d==1){
      pairedt$t1[i] <- myt
      pairedt$p1[i] <-myp
    }
    if(d==2){
      pairedt$t2[i] <-myt
      pairedt$p2[i] <-myp
    }
  }
}

# add t score to the data
t_dat <- select(pairedt, -3, -5)
# merge to allsum for analysis
allsum <- merge(allsum, t_dat)

cor(pairedt$t1,pairedt$t2,use='complete.obs')
```

@@

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
